\documentclass[12pt]{article}

\usepackage[]{amsmath}
\usepackage[]{amsthm}
\usepackage[]{amsfonts}
\usepackage[]{amssymb}
\usepackage{blindtext}
\usepackage{pgfplots}
\usepackage[a4paper, total={6in, 8in}]{geometry}

\title{Chapter 3:: Mathematical properties of state functions}

\author{Jerry Wu (217545898)}

\date{\today}
\begin{document}
\maketitle
\rule{\textwidth}{0.4pt}

\subsection*{Motivation}
We want to be able to express state functions as variables and partial derivatives. From there we can model stuff like internal energy and enthalpy in terms of $P$, $V$, and $P$ interchangably. We can also do the following::

\begin{itemize}
    \item Quantify the difference between $C_P$ and $C_V$
    \item Understand the Joule-Thomson experiment.
\end{itemize}

Consider $1mol$ of an ideal gas::
\begin{align*}
    P=f(V,T)=\frac{RT}{V}
\end{align*}

When we assume $T\equiv const$,
\begin{align*}
    \frac{\partial P}{\partial V}_T=\lim_{\Delta V\rightarrow 0}\frac{P(V+\Delta V,T)-P(V,T)}{\Delta V}=-\frac{RT}{V^2}
\end{align*}

When we assume $V\equiv const$,

\begin{align*}
    \frac{\partial P}{\partial T}_V=\lim_{\Delta V\rightarrow 0}\frac{P(V+\Delta V,T)-P(V,T)}{\Delta V}=\frac{R}{V}
\end{align*}

So the total change in pressure ($dP$) can be expressed as::

\begin{align*}
    dP=\frac{\partial P}{\partial T}_V dT+\frac{\partial P}{\partial V}_T dV
\end{align*}

In general, 

\begin{align*}
    dz=\frac{\partial z}{\partial x}_y dx + \frac{\partial z}{\partial y}_x dy
\end{align*}

by principle of vector addition.

\subsection*{The cyclic rule}

\begin{align*}
    \frac{\partial P}{\partial V}_T\frac{\partial V}{\partial T}_P\frac{\partial T}{\partial P}_V=-1\implies \frac{\partial P}{\partial T}_v=-\frac{\partial P}{\partial V}_T\frac{\partial V}{\partial T}_P=-\frac{\frac{\partial V}{\partial T}_P}{\frac{\partial V}{\partial P}}_T=\frac{\alpha}{\kappa_T}
\end{align*}

The expansion coefficient for isobaric volumetric thermal is::
\begin{align*}
    \alpha=\frac{1}{V}\frac{\partial V}{\partial T}_P
\end{align*}

Isothermal compressibility::

\begin{align*}
    \kappa_T=-\frac{1}{V}\frac{\partial V}{\partial P}_T
\end{align*}

So we have that

\begin{align*}
    dP=\frac{\alpha}{\kappa_T}dT-\frac{1}{\kappa_T V}dV
\end{align*}

and the total change is expressed as::

\begin{align*}
    \Delta P=\int_{T_i}^{T_f}\frac{\alpha}{\kappa_T}dT-\int_{V_i}^{V_f}\frac{1}{\kappa_T V}dV\approx \frac{\alpha}{\kappa_T}\Delta T-\frac{1}{\kappa_T}\ln(\frac{V_f}{V_i})
\end{align*}

\subsection*{Example 1}

You have accidentally arrived at the end of the range of an ethanol-in-glass
thermometer so that the entire volume of the glass capillary is filled. By how much
will the pressure in the capillary increase if the temperature is increased by another $10.0^{\circ}C$? $\alpha_{glass}=2E-5(^{\circ}C)^{-1}$,$\alpha_{ethanol}=11.2E-4(^{\circ}C)^{-1}$, $\kappa_{T,ethanol}=11.0E-5(bar)^{-1}$. Will the thermometer survive the experiment?

\begin{align*}
    \Delta P=\int_{T_i}^{T_f}\frac{\alpha_{ethanol}}{\kappa_{T,ethanol}}dT-\int_{V_i}^{V_f}\frac{1}{\kappa_{T,ethanol}V}dV\approx \frac{\alpha_{ethanol}}{\kappa_{T,ethanol}}(T_f-T_i)-\frac{1}{\kappa_{T,ethanol}}\ln(\frac{V_f}{V_i})
\end{align*}

So $\Delta P$ becomes::

\begin{align*}
    \Delta P=\frac{\alpha_{ethanol}}{\kappa_{T,ethanol}}\ln(\frac{V_i(1+\alpha_{glass}\Delta T)}{V_i})=100bar
\end{align*}

So no, the glass will not survive the experiment.

\subsection*{Dependence of $U$ on $V$ and $T$}

Internal energy on a fixed amount of substance is a function of $P$, $V$, and $T$.

\begin{align*}
    U=f(P_V)=f(V,T)=f(T,P)
\end{align*}

We are most interested in the one with volume and temperature.

\begin{align*}
    dU=\frac{\partial U}{\partial T}_V dT+\frac{\partial U}{\partial V}_T dV
\end{align*}

So how can numerical values for $\frac{\partial U}{\partial T}_V$ and $\frac{\partial U}{\partial V}_T$ be obtained?

\begin{align*}
    dq-P_{external}dV=\frac{\partial U}{\partial T}_V dT+\frac{\partial U}{\partial V}_T dV
\end{align*}

For an isochoric process, define the following::

\begin{align*}
    dq_V=\frac{\partial U}{\partial T}_V dT
\end{align*}
\begin{align*}
    \frac{dq_V}{dT}=\frac{\partial U}{\partial T}_V=C_V
\end{align*}

So now we can integrate for total change in internal energy.

\begin{align*}
    \Delta U_V=\int_{T_1}^{T_2}C_V dT=n\int_{T_1}^{T_2}C_{V,m}dT
\end{align*}

If $C_{V,m}$ is regarded as constant (small temperature range), we have that::

\begin{align*}
    \Delta U_V=\int_{T_1}^{T_2}C_V dT=C_V\Delta T=nC_{V,m}\Delta T
\end{align*}

\begin{align*}
    \int_{i}^{f}dq_V=\int_{i}^{f}\frac{\partial U}{\partial T}_V dT or q_V=\Delta U
\end{align*}

\subsection*{Internal pressure}

$\frac{\partial U}{\partial V}_T$ has units of $\frac{J}{m^3}$, which is internal pressure.

\begin{align*}
    \frac{\partial U}{\partial V}_T=T\frac{\partial P}{\partial T}_V-P
\end{align*}

Our total infinitessimal change in $U$ can be expressed as the following::

\begin{align*}
    dU=dU_V+dU_T=C_V dT+\left(T\frac{\partial P}{\partial T}_V-P\right) dV
\end{align*}

For $T\equiv const$

\begin{align*}
    dU=dU_T=\left(T\frac{\partial P}{\partial T}_V-P\right)dV
\end{align*}

For $V\equiv const$

\begin{align*}
    dU=dU_V=C_V dT
\end{align*}

So this means that $\Sigma dU=dU_V+dU_T$ by virtue of common sense not so common.

\subsection*{Example 2}

Evaluate $\frac{\partial U}{\partial V}_T$ for an ideal gas.

\begin{align*}
    \frac{\partial U}{\partial V}_T=T\frac{\partial P}{\partial T}_V-P=T\left(\frac{\partial \left( \frac{nRT}{V}\right)}{\partial T}\right)_V-P=\frac{nRT}{V}-P=0
\end{align*}

This is unsurprising, since ideal gases' $\Delta U$ don't depend on $\Delta V$. So we can conclude that $U$ is only a function of $T$ for ideal gases.\\We can observe that

\begin{align*}
    \Delta U_T=\int_{V_i}^{V_f}\frac{\partial U}{\partial V}_T dV\approx 0
\end{align*}

to a good approximation for $g\in \mathbb{R}$ under most conditions.\\ So what is the magnitude of this for processes involving liquids and solids?

\begin{align*}
    \Delta U_T^{solid,liquid}=\int_{V_1}^{V_2}\frac{\partial U}{\partial V}_T dV\approx \frac{\partial U}{\partial V}_T\Delta V\approx 0
\end{align*}

Under most conditions encountered in lab, $U$ can be regarded as a function of $T$ alone $\forall$ substances.

\begin{align*}
    U(T_f,V_f)-U(T_i,V_i)=\Delta U=\int_{T_i}^{T_f}C_V dT=n\int_{T_i}^{T_f}C_{V,m}dT
\end{align*}

\subsection*{Enthalpy as a state function}

$H$ is a function of $P$,$V$, and $T$.

\begin{align*}
    H=f(P,V)=f(V,T)=f(T,P)
\end{align*}

We can make $P\equiv const$.

After some work, we have that

\begin{align*}
    \int_{i}^{f}dU=\int_{i}^{f}dq_P-\int_{i}^{f}PdV
\end{align*}

\begin{align*}
    U_f-U_i=q_P-P(V_f-V_i)
\end{align*}

\subsection*{Phase chages}

For both fusion and vaporization (solid $\rightarrow$ liquid and liquid $\rightarrow$ gas), $q>0\implies C_P\rightarrow \infty$

\subsection*{Derivation}

Following our derivation of $U$ and its differentials, we can do the same for $H$.

\begin{align*}
    dH=\frac{\partial H}{\partial T}_P dT+\frac{\partial H}{\partial P}_T dP
\end{align*}

Because $dP=0$ at $P\equiv const$ and $dH=dq_P$, we have that

\begin{align*}
    dq_P=\frac{\partial H}{\partial T}_P dT
\end{align*}

and the heat capacity at $P\equiv const$ is defined as::

\begin{align*}
    C_P=\frac{dq_P}{dT}=\frac{\partial H}{\partial T}_P
\end{align*}

$C_P$ is determined by measuring the heat flow $dq$ at constant $P$ with the resulting temperature
change $dT$ in the limit in which $dT$ and $dq$ approach zero:

\begin{align*}
    C_P=\lim_{dT\rightarrow 0}\frac{dq}{dT}_P
\end{align*}

For a process where $P\equiv const$ assuming no phase changes or reactions, we have that

\begin{align*}
    \Delta H_P=\int_{T_i}^{T_f}C_P(T)dT=n\int_{T_i}^{T_f}C_{P,m}(T)dT
\end{align*}

\begin{align*}
    \Delta H_P=C_P\Delta T=nC_{P,m}\Delta T
\end{align*}

\subsection*{Example 3 (full solution on slide 30)}

A $143.0g$ sample of $C(s)$ in the form of graphite is heated from $300K \rightarrow 600K$ at a
constant pressure. Over this temperature range, $C_{P,m}$ has been determined to be

\begin{align*}
    \frac{C_{P,m}}{JK^{-1}mol^{-1}}=-12.19+0.1126\frac{T}{K}-(1.947E-4)\frac{T^2}{K^2}+(1.19E-7)\frac{T^3}{K^3}-(7.800E-11)\frac{T^4}{K^4}
\end{align*}

\begin{align*}
    \Delta H=\frac{m}{M}\int_{T_i}^{T_f}C_{P,m}(T)dT
\end{align*}

If $C_{P,m}$ at 300K is used instead, after some work, we have that $\Delta H=30.8kJ$

\end{document}