\documentclass[12pt]{book}

\usepackage[]{amsmath}
\usepackage[]{amsthm}
\usepackage[]{amsfonts}
\usepackage[]{amssymb}
\usepackage{blindtext}
\usepackage[a4paper, total={7.5in, 10.5in}]{geometry}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{array}
\usepackage{hyperref}


\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\pagenumbering{arabic}
\renewcommand{\chaptername}{Section}
\let\cleardoublepage\clearpage

\lstset{frame=tb,
  language=Java,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=left,
  numberstyle=\small\color{black},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=4
}

\title{HIST3801 Lecture 3}

\author{Jerry Wu}

\date{Sept 23 2024}

\begin{document}
\maketitle

\tableofcontents

\chapter{Lecture 3: History of Competitive Gaming}

\section{Spacewar}

\begin{itemize}
    \item Commonly known to be the first competitive gaming event. Developed by Steve Russell
    \item Limited access in high tech labs like MIT. Was only accessible by faculty or graduate students.
    \item Created on old radar screens from military technology; computer was bigger than a human (PDP1 minicomputer acquired by MIT in 1961)
    \item Paved the way towards titles like space invaders, but was originally a deathmatch game between 2 players
    \item The champion of the tournament was known as the "intergalactic spacewar champion". The winner was published in the Rolling Stone sports column and got a free \$25 subscription to said magazine for a year.
\end{itemize}

\section{Score keeping}

\begin{itemize}
    \item Scorekeeping is an old concept which has been used since ancient times to see which athlete was the superior competitor
    \item High scores first hit the market in 1976 with Midway's 1976 classic Seawolf
    \item High scores were scored on RAM, so as soon as the system gets turned off, the high score will get deleted. Machines had to be turned off at night to save money on electricity bills.
    \item By 1979, people were able to sign their initials beside their high scores on arcade machines
    \item The fact that high scores existed means that esports idols existed in newspapers and magazines. One could mail a photo of their high score to nintendo and they would feature it in a publication
    \item For example, Steve Wozniak was a god at tetris so he was at the top of the leaderboard consistently. He was dunking on 12 year old kids, so he was the first reddit mod before reddit existed.
    \item Billy Mitchell from Hollywood FL achieved a score of 1M in donkey kong and 3M in pac man. He was the very first god among gamers and was sponsored by people as a result which he used to found a restaurant franchise and hot sauce company
\end{itemize}

\subsection{Swordquest}

\begin{itemize}
    \item A tournament held by Atari where they were going to develop a series of fantasy games and people would beat them (speedrun) to find a code at the end of the game
    \item There were serious prizes for this tournament. The trophies for the tournament were made of real precious metals like gold and silver.
    \item Unfortunately the market was oversaturated and the project fell apart, so the prizes were just taken home by the organizers
\end{itemize}

\subsection{Besides high scores}

\begin{itemize}
    \item There exist other win conditions other than getting the highest score. For example, in a deathmatch game like Quake, the win condition is to eliminate all other players. This is an example of an alternate win condition
    \item The first quake tournament was in 1997, which was a LAN party at E3. The prize was a ferrari.
\end{itemize}

\subsection{The advent of the Internet, and the rise of esports proper}

\begin{itemize}
    \item By making things online, it opens doors to many possibilities for multiplayer without having to utilize LAN parties, saving transportation costs, living expenses, etc.
    \item LAN parties still remain popular for championships
    \item Games like starcraft, CS, league, etc all started spawning in this time period, and are still leaders in the market to this day
    \item Most households in Canada didn't have internet in the early 2000s, but this changed at around 2009
    \item Weird stuff started coming to fruition in this time period, such as video game porn (LOL)
\end{itemize}

\subsection{Are esports sports?}

\begin{itemize}
    \item There are many professional opinions on whether esports are sports or not. Some include:
    \begin{itemize}
        \item Sport has to be skill based and not luck based (true!)
        \item Sport has to involve physical movement (debatable)
        \item Sport has to be stable (regulated with rules, referees, tournaments, etc)
    \end{itemize}

    \item 
\end{itemize}

\end{document}